---
title: "HW12 Hints "
author: "COSC6323/Spring 2024"
date: "`r Sys.Date()`"
output: pdf_document
toc: true

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

```{r include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(cowplot)
library(ggpubr)
library(lme4)
library(readr)
library(lmerTest)
library(lattice)
library(gridExtra)
library(kableExtra)
require(ggpmisc) # for stat_poly_eq
library(bestNormalize)
library(FSA)
library(rstatix)
```

```{r include=FALSE}
rm(list = ls())
dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(dir)
getwd()
```



\newpage

# Question

We want to find out if there are any mean differences in terms of performance  among VIDEO, WORD, and ABSTRACT. Furthermore, if there are indeed differences, we would like to compute the relative order of sympathetic activation among the three treatments, that is, which one is the biggest, which one is in the middle, and which one is the lowest. 

Before applying the appropriate test, check if any pre-conditions are violated and take ameliorating measures as needed. 

\
\
\

# PLAN

  - Kruskal-Wallis test (one-way ANOVA)
  - Friedman test (repeated measures ANOVA)
  - Best normalization
  - Post-hoc analysis


1. The Kruskal-Wallis test when you have independent observations and non-normally distributed data across three or more groups. \

2. The Friedman test when dealing with repeated measures or ordinal(ranked) data \





```{r echo=FALSE}
Df <- read.csv("../Data/Affective-Math-Dataset_3.csv", stringsAsFactors = T)

# Exam data
Df_Exam <- Df[Df$Session == "Exam", ]
# names(Df_Exam)
# Remove unnecessary columns
Df_Exam <- subset(Df_Exam,
  select = c(
    ParticipantID, Perspiration, HR.E4, HR.AW,
    Question.Type, Question.Name, Time, Attempt, Accuracy.Score
  )
)


# Remove 2nd and 3rd attempts
Df_Exam <- Df_Exam[Df_Exam$Attempt == 1, ]

Df_Exam <- Df_Exam[Df_Exam$Question.Type != "Example", ]

# Remove NAs
Df_Exam <- Df_Exam[!is.na(Df_Exam$Question.Type), ]


Df_Exam <- droplevels(Df_Exam)
# str(Df_Exam)

# unique(Df_Exam$Question.Type)
# unique(Df_Exam$ParticipantID)
```


\newpage

```{r warning=FALSE, include=FALSE}
Qlevel <- Df_Exam %>%
  group_by(ParticipantID, Question.Name, Question.Type) %>%
  summarise(
    pp.mean = mean(Perspiration, na.rm = T),
    HR.E4.mean = mean(HR.E4, na.rm = T),
    HR.AW.mean = mean(HR.AW, na.rm = T),
    q.solv.time = n(),
    Accuracy.Score = first(Accuracy.Score)
  ) %>%
  convert_as_factor(Accuracy.Score)


# 1) For every subject compute the total sum of grades per question group

Qlevel2 <- Qlevel %>%
  group_by(ParticipantID, Question.Type) %>%
  mutate(
    n = n(),
    correct_score = sum(Accuracy.Score == 1),
  ) %>%
  convert_as_factor(ParticipantID, Question.Type)
```


# Summary statistics
```{r}
score.df <- Qlevel2 %>%
  select(ParticipantID, Question.Type, correct_score) %>%
  distinct() %>%
  convert_as_factor(ParticipantID, Question.Type) %>%
  ungroup()


score.df %>%
  group_by(Question.Type) %>%
  get_summary_stats(correct_score, type = "common") %>%
  kable("latex", booktabs = T, digits = 3) %>%
  kable_styling(latex_options = c("stripdown", "hold_position"), font_size = 7)
```

\
\
\

# Data visualization

```{r fig.height=5}
ggboxplot(score.df, x = "Question.Type", y = "correct_score", add = "jitter")


# score.df <- score.df[score.df$correct_score != 0,]

# table(score.df$correct_score)
# table(score.df$Question.Type)
# table(score.df$ParticipantID)

# colSums(is.na(score.df))
```

\newpage



# Kruskal-Wallis test (one-way ANOVA)

It is a non-parametric version of the one-way ANOVA,
situation where there are more than two groups. 

```{r}
# Kruskal-Wallis test

?kruskal_test()

result.kruskal <- kruskal_test(correct_score ~ Question.Type, data = score.df)
result.kruskal
```

Since the p-value is less than 0.05, we can reject the null hypothesis and conclude that there is a significant difference between the groups.


## Effect size of Kruskal-Wallis
```{r}
?rstatix::kruskal_effsize
score.df %>% kruskal_effsize(correct_score ~ Question.Type)

# The interpretation values commonly in published literature are: 
# 0.01- < 0.06 (small effect), 
# 0.06 - < 0.14 (moderate effect) 
# and >= 0.14 (large effect).
```

## Multiple pairwise-comparisons for Kruskal-Wallis

```{r}
?wilcox_test()
pwc.kw <- score.df %>%
  wilcox_test(correct_score ~ Question.Type,
    paired = FALSE,
    p.adjust.method = "bonferroni"
  )
# p.adjust.method = "bh")
# Adjusted p-values for multiple comparisons

pwc.kw
```

## Visualization of Kruskal-Wallis results

```{r}
pwc.kw <- pwc.kw %>% add_xy_position(x = "Question.Type")

ggboxplot(score.df, x = "Question.Type", y = "correct_score", add = "point") +
  stat_pvalue_manual(pwc.kw, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(result.kruskal, detailed = TRUE),
    caption = get_pwc_label(pwc.kw)
  )
```


## Dunn Test

```{r}
?dunnTest

# Pairwise comparisons
DT.score <- dunnTest(correct_score ~ Question.Type,
  data = score.df,
  # method="bh")
  method = "bonferroni"
)

# print(DT.score, dunn.test.results = TRUE)
DT.score
```

The pairwise comparison shows that, A and V groups are significantly different from each other.


\
\
\




\newpage
# Friedman test 

Similar to the parametric repeated measures ANOVA, 
it is used to detect if there are any statistically 
significant differences between the distributions 
of three or more paired groups.


```{r}
?friedman_test
res.fried <- score.df %>%
  friedman_test(correct_score ~ Question.Type | ParticipantID)

res.fried
```
Since the p-value is less than 0.05, we reject the null hypothesis that there is no difference in the mean scores among the three treatments.

Consider the following post-hoc tests to determine the relative order of sympathetic activation among the three treatments. 


## Effect size Friedman test
```{r}
?rstatix::friedman_effsize
score.df %>% friedman_effsize(correct_score ~ Question.Type | ParticipantID)

# The interpretation values commonly in published literature are:
# 0.01- < 0.06 (small effect), 
# 0.06 - < 0.14 (moderate effect) 
# and >= 0.14 (large effect).
```

## Multiple pairwise-comparisons

```{r}
# Multiple pairwise-comparisons for Kruskal-Wallis
pwc <- score.df %>%
  wilcox_test(correct_score ~ Question.Type, paired = TRUE, p.adjust.method = "bonferroni")

pwc
```


## Visualization of results
```{r}
pwc <- pwc %>% add_xy_position(x = "Question.Type")
ggboxplot(score.df, x = "Question.Type", y = "correct_score", add = "point") +
  stat_pvalue_manual(pwc, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res.fried, detailed = TRUE),
    caption = get_pwc_label(pwc)
  )
```


## Visualization of results with differences
```{r}

#Create different groups for the Question.Type variable


A = score.df[score.df$Question.Type =="A",]$correct_score
W = score.df[score.df$Question.Type =="W",]$correct_score
V = score.df[score.df$Question.Type =="V",]$correct_score

AW = A - W
AV = A - V
VW = V - W

diff.df = as.data.frame(cbind(AW, AV, VW))

boxplot(diff.df, col = c("red", "blue", "green"), main = "Boxplot of the differences")

```
A > V
W > V
A ? W


Results of the Friedman test indicate that there is a significant difference between the groups.




