---
title: "HW7 Solution "
author: "COSC6323/Spring 2024"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: true
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

----------

**Question 1: Multiple Linear Regression - HR vs. All Other Signals:** For the participants in the naturalistic dataset, perform multiple linear regression with HR as the response variable and predictors the following variables: Speed, Cadence, Altitude, Stride, Distances, Temperature, Power.  Justify your methodology and describe your insights.


\newpage

```{r include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(ggpubr)
library(lme4)
library(lmerTest) # get the best model
require(ggpmisc) # stat on plots with facets
library(rstatix) # tests
library(qqplotr)
library(sjPlot) # tab_model()
library(dplyr)

theme_set(theme_classic()) # set the theme to classic
theme_update(plot.title = element_text(hjust = 0.5)) # center the title

rm(list = ls())
dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(dir)
getwd()

```









```{r include=FALSE}


all.df2 = read.csv("../data/Multimodal-Activity-Dataset.csv", stringsAsFactors = T)

pace_split <- strsplit(as.character(all.df2$Pace), ":")
# Extracting hours and minutes using sapply and lapply
hours <- sapply(pace_split, function(x) as.numeric(x[1]))
minutes <- sapply(pace_split, function(x) as.numeric(x[2]) / 60)

# Calculating total pace in hours
all.df2$Pace2 <- hours + minutes


str(all.df2)

droplevels(all.df2)
str(all.df2$Pace)

# Check for missing values
colSums(is.na(all.df2))

```



```{r include=FALSE}

# all.df2$Time = as.POSIXct(all.df2$Time, format = "%Y-%m-%d %H:%M:%S")
# str(all.df2)


# Function to round timestamp to nearest 5 min
# round_timestamp <- function(timestamp) {
#   as.POSIXct(round(as.numeric(timestamp) / 10) * 10, origin = "1970-01-01")
# }
# 
# # create 5 min interval data set
# 
# all.df2_cut5min <- all.df2 %>%
#   group_by(ID, Activity) %>%
#   mutate(timestamp_rounded = round_timestamp(Time)) %>%
#   group_by(timestamp_rounded) %>%
#   summarise(across(where(~ !is.factor(.)), mean),
#             ID = first(ID),
#             Day = first(Activity),
#             Activity = first(Activity)) %>%
#   ungroup()
# 
# 
# all.df2 <- all.df2_cut5min



```

\newpage
# Question 1: Multiple Linear Regression - HR vs. All Other Signals

## Inspection of the data

```{r echo=TRUE}

browseURL("https://osf.io/m43vw/wiki/Homework%207/")
######### MODEL SOLUTION

# Fit the multiple linear regression model
lm.df = all.df2%>%
  dplyr::select(HR, Speed_NR, Cadence, Altitude, Distances, Temperatures, Power,Pace2) %>% drop_na()

str(lm.df)
sum(is.na(lm.df))
colSums(is.na(lm.df))

```

\newpage

# Multi Linear Model and diagnostic

## COLLINEARITY CHECK
```{r echo=TRUE}

######### COLLINEARITY CHECK
# Load the ggcorrplot package
library(ggcorrplot)
# Remove the HR column
reduced_data <- subset(lm.df, select = -c(HR))
colSums(is.na(reduced_data))


# Matrix scatterplot:
plot(reduced_data, pch=16, col="blue",
     main="Matrix Scatterplot of all precictors")



# Compute correlation at 2 decimal places
corr_matrix = round(cor(reduced_data), 2)


# Show the  result
#ggcorrplot(corr_matrix, hc.order = TRUE)#, type = "lower", lab = TRUE)
ggcorrplot(corr_matrix, type = "lower", lab = TRUE)



```





```{r echo=TRUE}
hr_multi_model <- lm(HR ~ Speed_NR + Cadence+ Altitude+ Distances +Temperatures + Power +Pace2, 
                data = lm.df)

summary(hr_multi_model)



```

```{r include=FALSE}
tab_model(hr_multi_model)
#save the model result as html
tab_model(hr_multi_model , file = "hr_cad_model.html")


# get coefficients
print(hr_multi_model)

# Get the Intercept and coefficients as vector elements.
coef(hr_multi_model)


# Criterions to identify the quality of the model
?AIC

#the smaller the AIC or BIC, the better the fit.

aic.full<- AIC(hr_multi_model)
cat("Full Model AIC:" ,aic.full)

```


```{r echo=TRUE, fig.height=7, fig.width=7}
# Diagnostic plots for the model
par(mfrow = c(2,2))
plot(hr_multi_model)

```


```{r echo=TRUE}

library(olsrr)
library(MASS)
library(stargazer)
library(dplyr)


step.optimal <- stepAIC(hr_multi_model, trace = TRUE, direction= "both")

# Compare the models
stargazer(hr_multi_model, step.optimal, type = "text")


# hr_multi_model %>% 
#   ols_step_backward_p() %>% plot()


```


```{r}

# Backward elimination
hr_multi_model %>% 
  ols_step_backward_p(details = TRUE)

# Forward elimination
hr_multi_model %>% 
  ols_step_forward_p(details = TRUE)

# Stepwise elimination
hr_multi_model %>%
  ols_step_both_p(details = TRUE)


```



\newpage
# REDUCED MODEL SOLUTION


## COLLINEARITY CHECK
```{r echo=TRUE}

# Remove the HR column
reduced_data <- subset(lm.df, select = -c(HR, Power))
colSums(is.na(reduced_data))

# Compute correlation at 2 decimal places
corr_matrix = round(cor(reduced_data), 2)

# Show the  result
#ggcorrplot(corr_matrix, hc.order = TRUE)#, type = "lower", lab = TRUE)
ggcorrplot(corr_matrix, type = "lower", lab = TRUE)

```




```{r echo=TRUE}
######### REDUCED MODEL SOLUTION
# Fit the reduced multiple linear regression model
hr_multi_model_reduced = lm(formula = HR ~ Speed_NR + Cadence + Altitude + Distances + Temperatures  + Pace2 , data = lm.df)

summary(hr_multi_model_reduced)


aic.reduced<- AIC(hr_multi_model_reduced)

cat("Reduced Model AIC:" ,aic.reduced)

cat("Full Model AIC:" ,aic.full)


```



# Compare Models

```{r echo=TRUE}
# https://www.scribbr.com/statistics/akaike-information-criterion/
#install.packages("AICcmodavg")
library(AICcmodavg)

# Compare the models
models <- list(hr_multi_model, hr_multi_model_reduced)

model.names <- c('hr_multi_model', 'hr_multi_model_reduced' )

comp = aictab(cand.set = models, modnames = model.names)
comp
```



