---
title: "HW13 Hints "
author: "COSC6323/Spring 2024"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: true
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```


\newpage

```{r include=FALSE}
library(tidyverse)
library(lubridate)
library(latex2exp)
library(cowplot)
library(ggpubr)
library(readr)
library(lattice)
library(gridExtra)
library(kableExtra)
library(lme4)
library(lmerTest) # get the best model
require(ggpmisc) # stat on plots with facets
library(rstatix) # tests
library(qqplotr)
library(sjPlot) # tab_model()
library(dplyr)
library(glmtoolbox) # Show the adjusted R2 in glm
library(arm) # convenience functions for regression in R display()
library(ggcorrplot)

theme_set(theme_classic()) # set the theme to classic
theme_update(plot.title = element_text(hjust = 0.5)) # center the title

rm(list = ls())
dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(dir)
getwd()
```




# Utility Functions
```{r echo=FALSE, fig.height=11, fig.width=20, message=FALSE, warning=FALSE}
# Without x axis label first rows
my_theme <- theme_bw() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 10, face = "bold"),
    strip.text = element_text(size = 18),
    legend.position = "none"
  )

# With  x-y axis label
my_theme1 <- my_theme +
  theme(
    axis.text.x = element_text(size = 8, face = "bold")
  )

activity_names <- c(
  `office_work` = "Office Work",
  `walking` = "Walking",
  `running` = "Running",
  `biking` = "Biking",
  `driving` = "Driving"
)

# helper for boxplot
n_fun <- function(x) {
  return(data.frame(y = median(x) * 1.3, label = paste0("~italic(n)", " == ", length(x))))
}


mycolors <- c(
  unknown = "white", office_work = "darkgray", walking = "darkolivegreen",
  running = "deepskyblue", biking = "orange",
  driving = "deeppink3"
)
```




# Read the raw data

```{r eval=FALSE, include=FALSE}
all.df2 <- read.csv(paste0("../Data/Naturalistic-Multimodal-Activity-Dataset_v5.csv"), stringsAsFactors = F)

colSums(is.na(all.df2))
str(all.df2)


all.df2$Time <- as.POSIXct(all.df2$Time, format = "%Y-%m-%d %H:%M:%S")
#
factor(all.df2$Day2)

all.df2$Day2 <- factor(all.df2$Day2,
  levels = c("PR", "MD", "MD2", "PS", "PS2", "RD", "RD2")
)
levels(all.df2$Period)

all.df2$Period <- as.factor(all.df2$Period)
all.df2$Gender <- as.factor(all.df2$Gender)

all.df2$Time2 <- format(all.df2$Time, "%H:%M:%S")

all.df2$Time2 <- as.POSIXct(all.df2$Time2, format = "%H:%M:%S")



# Check for missing values
colSums(is.na(all.df2))

str(all.df2)
```

## Cut the time into 1 min intervals

```{r eval=FALSE, include=FALSE}
colnames(all.df2)

# Add Day2 Label here
# Remove NAs from the data
str(all.df2)
colnames(all.df2)
colSums(is.na(all.df2))


# helper to get the max values names from the aggreegated data.
max_freq <- function(x) {
  return(names(which.max(table(x))))
}


all.df2_cut <- all.df2 %>%
  # filter(ID == "T009") %>%
  group_by(ID, Time) %>%
  # Time needs to be Posixct
  mutate(
    Time =
      as.POSIXct(Time, format = "%Y-%m-%d %H:%M:%S") %>%
        floor_date("1 min")
  ) %>%
  summarise(
    ID = first(ID),
    Day = first(Day),
    Activity = first(Activity),
    Day2 = first(Day2),
    Period = first(Period),
    # return name of max index
    max_ind = max_freq(max_ind),
    HR_normalized = mean(HR_normalized, na.rm = TRUE),
    HR = mean(HR, na.rm = TRUE),
    Speed = mean(Speed, na.rm = TRUE),
    Cadence = mean(Cadence, na.rm = TRUE),
    Altitude = mean(Altitude, na.rm = TRUE),
    Distances = mean(Distances, na.rm = TRUE),
    Temperatures = mean(Temperatures, na.rm = TRUE),
    Power = mean(Power, na.rm = TRUE),
    Speed_NR = mean(Speed_NR, na.rm = TRUE),
    Pace2 = mean(Pace2, na.rm = TRUE),
    # new variables
    Age = first(Age),
    Gender = first(Gender),
    BMI = first(BMI)
  ) %>%
  ungroup()

# save
write.csv(all.df2_cut, "Naturalistic-Multimodal-Activity-Dataset_v5_1min.csv", row.names = F)


```

## Read aggregated data

```{r include=FALSE}
all.df2_cut = read.csv("Naturalistic-Multimodal-Activity-Dataset_v5_1min.csv", stringsAsFactors = T )

all.df2_cut$Time <- as.POSIXct(all.df2_cut$Time, format = "%Y-%m-%d %H:%M:%S")

all.df2_cut$Time2 <- format(all.df2_cut$Time, "%H:%M:%S")
all.df2_cut$Time2 <- as.POSIXct(all.df2_cut$Time2, format = "%H:%M:%S")

all.df2_cut$Day2 <- factor(all.df2_cut$Day2,
  levels = c("PR", "MD", "MD2", "PS", "PS2", "RD")
)

# Have only 4 levels for Day2 data
all.df2_cut$Day3 <- recode_factor(all.df2_cut$Day2,
                                  "PR" = "PR", "PR2" = "PR", "MD" = "MD", "MD2" = "MD", "PS" = "PS", "PS2" = "PS", "RD" = "RD"
)

all.df2_cut$Day2 <- all.df2_cut$Day3 %>% as.factor()
all.df2_cut$Day3 <- NULL

# Day2 reference level is MD
all.df2_cut$Day2 <- relevel(all.df2_cut$Day2, ref = "MD")
colSums(is.na(all.df2_cut))

# Period reference level is Afternoon
all.df2_cut$Period <- relevel(all.df2_cut$Period, ref = "Afternoon")
colSums(is.na(all.df2_cut))

# Activity reference level is office_work
all.df2_cut$max_ind <- relevel(all.df2_cut$max_ind, ref = "office_work")
colSums(is.na(all.df2_cut))

all.df3_cut <- all.df2_cut %>%
  rename(
  Participant = ID,
  Init.Day = Day,
  Day = Day2,
  Distance = Distances,
  Init.Speed = Speed,
  Speed = Speed_NR,
  Temperature = Temperatures,
  Init.Activity = Activity,
  Activity = max_ind
  )

levels(all.df3_cut$Activity) <- c("Office.Home", "Biking", "Driving", "Running", "Unknown", "Walking")


# Remove NAs from the data
all.df3_cut <- all.df3_cut %>%
  filter(Day != "NA") %>%
  filter(Activity != "Unknown") %>% droplevels()


all.df3_cut$Activity

```



## Create boxplot for HR by Day

```{r echo=FALSE}
all.df3_cut %>%
  ggplot(aes(x = Day, y = HR, fill = Day)) +
  geom_boxplot() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Boxplot of HR by Day",
    x = "Day",
    y = "HR"
  ) +   stat_summary(
    fun.data = n_fun, geom = "text", fun.y = median, angle = 90,
    parse = T, size = 3, vjust = -0.75, hjust = 0,
    position = position_dodge(width = 0.75)
  )
```


\newpage
# Question 1 - Generalized Linear Regression - HR vs. Various Predictors with Fixed Effects


```{r include=FALSE}
## Inspection of the data

all.df.glm <- all.df3_cut %>%
  dplyr::select(Participant, HR, Speed, Cadence, Altitude, Distance, Temperature, Power, Pace2, Period, Age, BMI, Activity, Day) %>%
  drop_na()

dim(all.df.glm)
str(all.df.glm)
sum(is.na(all.df.glm))
colSums(is.na(all.df.glm))
```


# Multi Linear Model and diagnostic

## COLLINEARITY CHECK including all predictors

```{r fig.height=10, fig.width=10, warning=FALSE}

reduced_data <- subset(all.df.glm, select = -c(Participant))

model.matrix(~ 0 + ., data = reduced_data) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(show.diag = FALSE, type = "lower", lab = TRUE)
```







<!-- Ref.:  -->
<!-- https://rpubs.com/JSHAH/481706 -->
<!-- https://medium.com/analytics-vidhya/interpreting-acf-or-auto-correlation-plot-d12e9051cd14 -->
<!-- https://towardsdatascience.com/identifying-ar-and-ma-terms-using-acf-and-pacf-plots-in-time-series-forecasting-ccb9fd073db8 -->
<!-- https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html -->




```{r echo=FALSE}

library(tseries)
library(forecast)

HR_data = ts(all.df.glm$HR)

is.ts(HR_data)

summary(HR_data)
start(HR_data)
end(HR_data)
time(HR_data)
frequency(HR_data)
ts.plot(HR_data)
acf(HR_data)
pacf(HR_data)

```


```{r echo=FALSE}
# Auto correlation

?acf
acf(HR_data,  plot = F)
acf(HR_data, plot = TRUE)
acf(HR_data, lag.max = 720, plot = T)

HRts = ts(HR_data, frequency = 1) # Since the data big, 1 is used to save time
#HRts = ts(HR_data, frequency = 10)
#HRts = ts(HR_data, frequency = 20)
#HRts = ts(HR_data, frequency = 60)


#partial auto correlation
#stationary test
?adf.test() # check our ts is stationary or not
adf.test(HR_data, alternative = "stationary", k = 1)
adf.test(HR_data, alternative = "stationary", k = 4)
adf.test(HR_data, alternative = "stationary", k = 720)



```

# Find Best ARIMA model


```{r echo=FALSE}
# since P values is less than 0.05, we can say that our data is stationary that means we can use ARIMA model

find_best = auto.arima(HRts, 
                       stepwise = FALSE,
                       approximation = FALSE, 
                       trace = F, ic = "aicc")
find_best


```
```{r}
HRts %>%
  Arima(order=c(1,1,4), seasonal=c(1,1,4)) %>%
  residuals() %>% ggtsdisplay()
```



```{r echo=FALSE}

# check residuals
qqnorm(find_best$residuals)
qqline(find_best$residuals)
hist(find_best$residuals)


```

# Forecast the data
```{r echo=FALSE}
# forecast the data
plot(forecast(find_best, h = 300))

```


- Autocorrelation represents the degree of similarity between a given time series 
and a lagged version of itself over successive time intervals.
- Autocorrelation measures the relationship between a variable’s current value and its past values.
- An autocorrelation of +1 represents a perfect positive correlation, 
while an autocorrelation of negative 1 represents a perfect negative correlation.


*Now time pay attention more!*


## HR Final Model | Best model from HW9


```{r echo=FALSE}

hr_multi_model <-
  glmer(
  formula = HR ~ Cadence + Altitude + Temperature + Day*Period + Activity + (1 | Participant),
  data = all.df.glm
)

summary(hr_multi_model)


acf(residuals(hr_multi_model))
pacf(residuals(hr_multi_model))

#plot(residuals(hr_multi_model))


sjPlot::plot_model(hr_multi_model,
                   show.values=TRUE, show.p=TRUE,
                   title="") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
          axis.text = element_text(size = 10)
       ) + theme_bw()


```


\newpage

## HR Final Model - Diagnostic ACF and PACF

```{r echo=FALSE}

HR = all.df.glm$HR
Participant = all.df.glm$Participant
Speed = all.df.glm$Speed
Cadence = all.df.glm$Cadence
Altitude = all.df.glm$Altitude
Temperature = all.df.glm$Temperature
Power = all.df.glm$Power
Period = all.df.glm$Period
Activity = all.df.glm$Activity
Day = all.df.glm$Day


HR1<-HR[2:length(HR)]
HR1lag1<-HR[1:(length(HR)-1)]

### Random Effects in the model:
Participant1<-Participant[2:length(HR)]

### Fixed Effects in the model:
Speed1<-Speed[2:length(HR)]
Cadence1<-Cadence[2:length(HR)]
Altitude1<-Altitude[2:length(HR)]
Temperature1<-Temperature[2:length(HR)]
Power1<-Power[2:length(HR)]
Period1<-Period[2:length(HR)]
Activity1<-Activity[2:length(HR)]
Day1<-as.factor(Day[2:length(HR)])

### Create a data frame with the variables

hr_multi_model2 = glmer(HR1 ~ HR1lag1 + (1|Participant1)  + Cadence1 + Altitude1 + Temperature1  + Period1 + Activity1 + Day1)

summary(hr_multi_model2)

acf(residuals(hr_multi_model2))
pacf(residuals(hr_multi_model2))

sjPlot::plot_model(hr_multi_model2,
                   show.values=TRUE, show.p=TRUE,
                   title="") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
          axis.text = element_text(size = 10)
       ) + theme_bw()




```

```{r echo=FALSE}
hr_multi_model3 = glmer(HR1 ~ HR1lag1 + (1|Participant1)  + Cadence1 + Altitude1 + Temperature1  + Day1*Period1 + Activity1 )


summary(hr_multi_model3)

acf(residuals(hr_multi_model3))
pacf(residuals(hr_multi_model3))

sjPlot::plot_model(hr_multi_model3,
                   show.values=TRUE, show.p=TRUE,
                   title="") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
          axis.text = element_text(size = 10)
       ) + theme_bw()

```

