---
title: "HW8 Hints "
author: "COSC6323/Spring 2024"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: true
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

----------

**Question 1: **


\newpage

```{r include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(ggpubr)
library(lme4)
library(lmerTest) # get the best model
require(ggpmisc) # stat on plots with facets
library(rstatix) # tests
library(qqplotr)
library(sjPlot) # tab_model()
library(dplyr)

theme_set(theme_classic()) # set the theme to classic
theme_update(plot.title = element_text(hjust = 0.5)) # center the title

rm(list = ls())
dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(dir)
getwd()

```




# Utility Functions
```{r echo=FALSE, fig.height=11, fig.width=20, message=FALSE, warning=FALSE}
# Without x axis label first rows
my_theme <- theme_bw() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 10, face = "bold"),
    strip.text = element_text(size = 18),
    legend.position = "none"
  )

# With  x-y axis label
my_theme1 <- my_theme +
  theme(
    axis.text.x = element_text(size = 8, face = "bold")
  )

activity_names <- c(
  `office_work` = "Office Work",
  `walking` = "Walking",
  `running` = "Running",
  `biking` = "Biking",
  `driving` = "Driving"
)

mycolors <- c(
  unknown = "white", office_work = "darkgray", walking = "darkolivegreen",
  running = "deepskyblue", biking = "orange",
  driving = "deeppink3"
)
```


# Read Reference Data
```{r include=FALSE}
# Read the all.df file
dataDir <- paste0("../Data/")
file <- "Multimodal-Activity-Dataset.csv"
all.df <- read.csv(paste0(dataDir, file), stringsAsFactors = F)
str(all.df)

unique(all.df$ID)
summary(all.df)

nlevels(all.df$Activity)
unique(all.df$Activity)
all.df$Activity <- factor(all.df$Activity, levels = c("office_work", "walking", "running", "biking", "driving"))
nlevels(all.df$Activity)
unique(all.df$Activity)

# Find the NA rows in HR
all.df.HR.NAs <- all.df[is.na(all.df$HR), ]
nrow(all.df.HR.NAs)

# Remove all data points where Cadence is bigger than 0 in driving
outliers_Cad_Dr <- all.df[all.df$Cadence > 0 & all.df$Activity == "driving", ]
# Remove outliers
all.df <- all.df %>% anti_join(outliers_Cad_Dr)

# Remove the NA rows in HR
all.df <- all.df[!is.na(all.df$HR), ]
all.df <- all.df %>% group_by(ID)



## Remove data point where HR is NA, Most of the case any HR file data has NA
activity_names <- c(
  `office_work` = "Office Work",
  `walking` = "Walking",
  `running` = "Running",
  `biking` = "Biking",
  `driving` = "Driving"
)

# Testing density function
tmp.d <- density(all.df[all.df$Activity == "office_work", ]$HR)
plot(tmp.d, main = "Density Estimation")
tmm.p.d_fun <- approxfun(tmp.d$x, tmp.d$y)

tmm.p.d_fun <- approxfun(density(all.df[all.df$Activity == "office_work", ]$HR))
tmm.p.d_fun(75)

activity <- "office_work"

# Channel based functions
density_fun_HR <- function(x, activity) {
  all.df.HR <- density(all.df[all.df$Activity == activity, ]$HR)
  all.df.HR_fun <- approxfun(all.df.HR$x, all.df.HR$y)
  return(all.df.HR_fun(x))
}

density_fun_HR(0:200, activity)

density_fun_Speed <- function(x, activity) {
  all.df.Speed_NR <- density(all.df[all.df$Activity == activity, ]$Speed_NR)
  all.df.Speed_fun <- approxfun(all.df.Speed_NR$x, all.df.Speed_NR$y)
  return(all.df.Speed_fun(x))
}

density_fun_Speed(0:75, activity)

density_fun_Cadence <- function(x, activity) {
  all.df.Cadence <- density(all.df[all.df$Activity == activity, ]$Cadence)
  all.df.Cadence_fun <- approxfun(all.df.Cadence$x, all.df.Cadence$y)
  return(all.df.Cadence_fun(x))
}

all.df.Cadence <- density(all.df[all.df$Activity == activity, ]$Cadence)
plot(all.df.Cadence, main = "Density Estimation")
all.df.Cadence_fun <- approxfun(all.df.Cadence$x, all.df.Cadence$y)

density_fun_Cadence(1:100, activity)

# Create a function for each channel and activity

# HR
density_fun_HR_office_work <- function(x) density_fun_HR(x, "office_work")
density_fun_HR_walking <- function(x) density_fun_HR(x, "walking")
density_fun_HR_running <- function(x) density_fun_HR(x, "running")
density_fun_HR_biking <- function(x) density_fun_HR(x, "biking")
density_fun_HR_driving <- function(x) density_fun_HR(x, "driving")

density_fun_HR_biking(40:200)

# Speed
density_fun_Speed_office_work <- function(x) density_fun_Speed(x, "office_work")
density_fun_Speed_walking <- function(x) density_fun_Speed(x, "walking")
density_fun_Speed_running <- function(x) density_fun_Speed(x, "running")
density_fun_Speed_biking <- function(x) density_fun_Speed(x, "biking")
density_fun_Speed_driving <- function(x) density_fun_Speed(x, "driving")

density_fun_Speed_office_work(10)

# Cadence
density_fun_Cadence_office_work <- function(x) density_fun_Cadence(x, "office_work")
density_fun_Cadence_walking <- function(x) density_fun_Cadence(x, "walking")
density_fun_Cadence_running <- function(x) density_fun_Cadence(x, "running")
density_fun_Cadence_biking <- function(x) density_fun_Cadence(x, "biking")
density_fun_Cadence_driving <- function(x) density_fun_Cadence(x, "driving")

density_fun_Cadence_office_work(10)

##### ------ Channel based functions --- end
```


```{r message=FALSE, warning=FALSE, include=FALSE}
# My density function
channel <- "HR"
activity <- "driving"

rm(density_fun)
density_fun <- function(input, activity, channel) {
  tmp.density <- density(all.df[all.df$Activity == activity, c(channel)][[1]])
  tmp.density_fun <- approxfun(tmp.density$x, tmp.density$y)
  return(tmp.density_fun(input))
}
```



# Read the raw data

```{r echo=TRUE}


# Read 6 subjects data


all.df2 = read.csv(paste0("../data/Naturalistic-Multimodal-Activity-Dataset_v4.csv"), stringsAsFactors = F)

colSums(is.na(all.df2))
str(all.df2)
all.df2 <- all.df2 %>% drop_na(Time)




all.df2$Time <- as.POSIXct(all.df2$Time, format = "%Y-%m-%d %H:%M:%S")
# 
#factor(all.df2$Day2)
 
all.df2$Day2 <- factor(all.df2$Day2 , 
                        levels = c("PR", "PR2", "MD", "MD2",  "PS", "PS2", "RD", "RD2"))
levels(all.df2$Period)

all.df2$Period <- as.factor(all.df2$Period)
all.df2$Gender <- as.factor(all.df2$Gender)

all.df2$Time2 <- format(all.df2$Time, "%H:%M:%S")

all.df2$Time2 <- as.POSIXct(all.df2$Time2, format = "%H:%M:%S")



# Check for missing values
colSums(is.na(all.df2))

str(all.df2)


```



## PDF implementation

```{r include=FALSE}
unique(all.df2$ID)
table(unique(all.df2$Day2))

all.df2 <- droplevels(all.df2)
colnames(all.df2)

all.df_extented_3 <- all.df2 %>%
  mutate(
    Offc_HR.dens = density_fun_HR_office_work(HR),
    Offc_Speed_NR.den = density_fun_Speed_office_work(Speed_NR),
    Offc_Cadence.dens = density_fun_Cadence_office_work(Cadence)
  ) %>%
  mutate(
    Wlkg_HR.dens = density_fun_HR_walking(HR),
    Wlkg_Speed_NR.den = density_fun_Speed_walking(Speed_NR),
    Wlkg_Cadence.dens = density_fun_Cadence_walking(Cadence)
  ) %>%
  mutate(
    Rnng_HR.dens = density_fun_HR_running(HR),
    Rnng_Speed_NR.den = density_fun_Speed_running(Speed_NR),
    Rnng_Cadence.dens = density_fun_Cadence_running(Cadence)
  ) %>%
  mutate(
    Bkg_HR.dens = density_fun_HR_biking(HR),
    Bkg_Speed_NR.den = density_fun_Speed_biking(Speed_NR),
    Bkg_Cadence.dens = density_fun_Cadence_biking(Cadence)
  ) %>%
  mutate(
    Drvg_HR.dens = density_fun_HR_driving(HR),
    Drvg_Speed_NR.den = density_fun_Speed_driving(Speed_NR),
    Drvg_Cadence.dens = density_fun_Cadence_driving(Cadence)
  ) %>%
  mutate(
    Offc_multi.dens = Offc_HR.dens * Offc_Speed_NR.den * Offc_Cadence.dens,
    Wlkg_multi.dens = Wlkg_HR.dens * Wlkg_Speed_NR.den * Wlkg_Cadence.dens,
    Rnng_multi.dens = Rnng_HR.dens * Rnng_Speed_NR.den * Rnng_Cadence.dens,
    Bkg_multi.dens = Bkg_HR.dens * Bkg_Speed_NR.den * Bkg_Cadence.dens,
    Drvg_multi.dens = Drvg_HR.dens * Drvg_Speed_NR.den * Drvg_Cadence.dens
  )
```


```{r include=FALSE}
# Replace NA with 0 in probability products values.
# Replace values less than 0.000001 with 0
colnames(all.df_extented_3)

denscols = c("Offc_HR.dens",
             "Offc_Speed_NR.den",
             "Offc_Cadence.dens", "Wlkg_HR.dens", "Wlkg_Speed_NR.den", "Wlkg_Cadence.dens",
             "Rnng_HR.dens", "Rnng_Speed_NR.den", "Rnng_Cadence.dens",
             "Bkg_HR.dens", "Bkg_Speed_NR.den", "Bkg_Cadence.dens",
             "Drvg_HR.dens", "Drvg_Speed_NR.den", "Drvg_Cadence.dens" )

all.df_extented_3[, denscols] %>%
  replace(., is.na(.), 0) -> all.df_extented_3[, denscols]

all.df_extented_3[, denscols] %>%
  replace(., . < 0.000001, 0) -> all.df_extented_3[, denscols]


# Replace NA with 0 in probability products values.
all.df_extented_3 <- all.df_extented_3 %>%
  mutate_at(vars(Offc_multi.dens, Wlkg_multi.dens, Rnng_multi.dens, Bkg_multi.dens, Drvg_multi.dens), ~ replace_na(., 0))


# Find the max value in each row
all.df_extented_3$max_indVal <- apply(all.df_extented_3[, c(
  "Offc_multi.dens", "Wlkg_multi.dens",
  "Rnng_multi.dens", "Bkg_multi.dens",
  "Drvg_multi.dens"
)], 1, max)

# Find the column name of the max value in each row if not all of them are zero
all.df_extented_3 %>%
  mutate(
    max_ind = pmap_dbl(list(
      Offc_multi.dens, Wlkg_multi.dens, Rnng_multi.dens, Bkg_multi.dens,
      Drvg_multi.dens ), ~ which.max(c(...)))
  ) -> all.df_extented_3



# If all products are zero (max_indVal = 0), then the max_ind should be zero
all.df_extented_3[all.df_extented_3$max_indVal == 0, ]$max_ind <- 0

all.df_extented_3$max_ind <- as.factor(all.df_extented_3$max_ind)

# Recode the factor
all.df_extented_3$max_ind <- recode_factor(all.df_extented_3$max_ind,
  `0` = "unknown", `1` = "office_work",
  `2` = "walking", `3` = "running", `4` = "biking",
  `5` = "driving"
)


str(all.df_extented_3)
table(all.df_extented_3$ID)
```


## Cut the time into 5 min intervals
```{r include=FALSE}
colnames(all.df_extented_3)
all.df_extented_3_2 <- all.df_extented_3[,c(1:23,45)]
colnames(all.df_extented_3_2)


# Add Day2 Label here
# Remove NAs from the data
str(all.df_extented_3_2)
colnames(all.df_extented_3_2)
colSums(is.na(all.df_extented_3_2))


# helper to get the max values names from the aggreegated data.
max_freq <- function(x) {
  return(names(which.max(table(x))))
}


all.df2_cut <- all.df_extented_3_2 %>%
  #filter(ID == "T009") %>%
  group_by(ID,Time) %>%
  # Time needs to be Posixct
  mutate(Time =
           as.POSIXct(Time, format = "%Y-%m-%d %H:%M:%S") %>%
           floor_date("5 min")) %>%
  summarise(ID = first(ID),
            Day = first(Day),
            Activity = first(Activity),
            Day2 = first(Day2),
            Period = first(Period),
            # return name of max index
            max_ind = max_freq(max_ind),
            HR_normalized = mean(HR_normalized, na.rm = TRUE),
            HR = mean(HR, na.rm = TRUE),
            Speed = mean(Speed, na.rm = TRUE),
            Cadence = mean(Cadence, na.rm = TRUE),
            Altitude = mean(Altitude, na.rm = TRUE),
            Distances = mean(Distances, na.rm = TRUE),
            Temperatures = mean(Temperatures, na.rm = TRUE),
            Power = mean(Power, na.rm = TRUE),
            Speed_NR = mean(Speed_NR, na.rm = TRUE),
            Pace2 = mean(Pace2, na.rm = TRUE),
            # new variables
            Age = first(Age),
            Gender = first(Gender),
            BMI = first(BMI)
            
            ) %>%
  ungroup()



levels(all.df2_cut$Day2)

all.df2_cut$Day2 <- factor(all.df2_cut$Day2 , 
                                 levels = c("PR", "PR2", "MD", "MD2",  "PS", "PS2", "RD")
                                 )

#all.df2_cut$Time <- as.POSIXct(all.df2_cut$Time, format = "%Y-%m-%d %H:%M:%S")
all.df2_cut$Time2 <- format(all.df2_cut$Time, "%H:%M:%S")
all.df2_cut$Time2 <- as.POSIXct(all.df2_cut$Time2, format = "%H:%M:%S")
colnames(all.df2_cut)

# Recode the factor
all.df2_cut$max_ind <- recode_factor(all.df2_cut$max_ind,
  `0` = "unknown", `1` = "office_work",
  `2` = "walking", `3` = "running", `4` = "biking",
  `5` = "driving"
)


str(all.df2_cut)


```




\newpage
# Question 1




## Inspection of the data

```{r echo=TRUE}

browseURL("https://osf.io/m43vw/wiki/Homework%208/")
######### MODEL SOLUTION
# Fit the multiple linear regression model

all.df.glm = all.df2_cut%>%
  #filter(Day2 == "MD") %>%
  filter(max_ind != "unknown") %>%
  #dplyr::select(ID, Day2,HR, Speed_NR, Cadence, Altitude, Distances, Temperatures, Power,Pace2, Age, Gender, BMI) %>% 
  dplyr::select(HR, Speed_NR, Cadence, Altitude, Distances, Temperatures, Power, Pace2, Period, Age, BMI, max_ind) %>% 
  drop_na()

dim(all.df.glm)
str(all.df.glm)
sum(is.na(all.df.glm))
colSums(is.na(all.df.glm))



```

\newpage

# Multi Linear Model and diagnostic

## COLLINEARITY CHECK
```{r echo=TRUE}

######### COLLINEARITY CHECK
library(ggcorrplot)
# Remove the HR column
reduced_data <- subset(all.df.glm, select = -c(HR, Age, Period, max_ind))
colSums(is.na(reduced_data))
str(reduced_data)

# Compute correlation at 2 decimal places
corr_matrix = round(cor(reduced_data), 2)

# Show the  result for predictors
#ggcorrplot(corr_matrix, hc.order = TRUE)#, type = "lower", lab = TRUE)
ggcorrplot(corr_matrix, type = "lower", lab = TRUE)

```


# HR Full Model 

```{r echo=TRUE}
?lm()
?glm()

hr_multi_model <- glm(HR ~ Speed_NR + Cadence+ Altitude+ Distances +Temperatures + Power +Pace2 + Period + Age + BMI + max_ind, data = all.df.glm)

summary(hr_multi_model)

hr_multi_model <- glm(HR ~ Cadence+ Altitude+ Distances +Temperatures + Power +Pace2 + Period + Age + BMI + max_ind, data = all.df.glm)


hr_multi_model <- glm(HR ~ 1, data = all.df.glm) # forward starts
summary(hr_multi_model)


```

# HR Full Model plots
```{r}
m.plot <- sjPlot::plot_model(hr_multi_model,
                   show.values=TRUE, show.p=TRUE,
                   title="") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
          axis.text = element_text(size = 10)
       ) + theme_bw()

#m.plot$data$term
m.plot

```

## Full Model Assumption Checking


```{r echo=TRUE, fig.height=7, fig.width=7}
par(mfrow = c(2,2))
plot(hr_multi_model)

```


```{r echo=TRUE}

library(olsrr)
library(MASS)
library(stargazer)
library(dplyr)


step.optimal <- stepAIC(hr_multi_model, trace = TRUE, direction= "both")

# Compare the models
stargazer(hr_multi_model, step.optimal, type = "text")


```


```{r eval=FALSE, include=FALSE}

# Backward elimination
optimal_be = hr_multi_model %>% 
  ols_step_backward_p(details = F)

optimal_be

# # Forward elimination
# optimal_fe = hr_multi_model %>% 
#   ols_step_forward_p(details = TRUE)
# 
# # Stepwise elimination
# optimal_se = hr_multi_model %>%
#   ols_step_both_p(details = TRUE)

```


# Backward Elimination (BE) Optimal Model Selection

```{r echo=TRUE}

step_bw <- step(hr_multi_model, direction = "backward", trace = TRUE)
summary(step_bw) # Display elimination results


```
## get model from step_bw

```{r}
#get model from step_bw
final_hr_multi_model.bw <- eval(step_bw$call)
summary(final_hr_multi_model.bw)
AIC(final_hr_multi_model.bw)

```

## Final Model Assumption Checking

```{r echo=FALSE, fig.height=7}
par(mfrow=c(2,2))
plot(final_hr_multi_model.bw)
```

# Final HR Model plots

```{r}
m.plot <- sjPlot::plot_model(final_hr_multi_model.bw,
                   show.values=TRUE, show.p=TRUE,
                   title="") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
          axis.text = element_text(size = 10)
       ) + theme_bw()

m.plot
```


## Final Model Predictors Plots

```{r}
summary(final_hr_multi_model.bw)
activity.plot <-plot_model(final_hr_multi_model.bw, 
             type = "pred", 
             terms = "max_ind",
             dot.size = 5,
             line.size = 2,
             title = "Activity") +  theme_bw()+

    aes(color = c("1", "2", "3", "4", "5")) +
    #scale_color_manual( values = c( "gray","orange","orange")) +
      theme(panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        legend.position = "none"
        )+ labs(x="", y="")


summary(final_hr_multi_model.bw)
period.plot <-plot_model(final_hr_multi_model.bw, 
             type = "pred", 
             terms = "Period",
             dot.size = 6,
             line.size = 2,
             title = "Period") +  theme_bw()+
      theme(panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(size = 10, face = "bold"),
        legend.position = "none"
        )+ labs(x="", y="")


```


```{r fig.height=5, fig.width=10, warning=FALSE}
ggarrange(activity.plot, period.plot, ncol = 2, nrow = 1)

```


#  HR vs. Various Predictors with Mixed Effects

```{r}

all.df.glm = all.df2_cut%>%
  filter(Day2 == "MD") %>%
  filter(max_ind != "unknown") %>%
  #dplyr::select(ID, Day2,HR, Speed_NR, Cadence, Altitude, Distances, Temperatures, Power,Pace2, Age, Gender, BMI) %>% 
  dplyr::select(ID,HR, Speed_NR, Cadence, Altitude, Distances, Temperatures, Power, Pace2, Period, Age, BMI, max_ind) %>% 
  drop_na()


hr_multi_model <- lmer(HR ~ Speed_NR + Cadence+ Altitude+ Distances +Temperatures + Power +Pace2  + Period + Age + BMI + max_ind + (1 | ID), data = all.df.glm)

summary(hr_multi_model)

```

# HR Final Model plots
```{r}
step_bw <- step(hr_multi_model, direction = "backward", trace = TRUE)
step_bw # Display elimination results

```
## get model from step_bw

```{r}
#get model from step_bw
final_hr_multi_model.bw <- lmer(formula = HR ~ Cadence + Distances + Temperatures + Power + 
    Pace2 + Period + Age + BMI+ max_ind + (1+ID), data = all.df.glm)


summary(final_hr_multi_model.bw)


```



